---
title: "conos"
output: html_notebook
---

Author: Katarina Dragicevic (PhD student), Biotech Research and Innovation Centre, University of Copenhagen

We take the cleaned Conos object from CRMetrics output. This means, once we filtered out the cells, we run more time the processing in CRMetrics pipeline to get a conos object and a UMAP with clan data. Then we use that object.

```{r}
library(CRMetrics)
library(conos)

con <- crm$con
con$findCommunities(min.group.size = 100)
con$plotGraph()
```


We can also get the same by running the full conos pipeline on the normalized count matrices:

```{r}
source("http://pklab.med.harvard.edu/rasmus/scRNA_helper.R")

library(conos)
library(pagoda2)
con <- Conos$new(panel, n.cores=200)

con <- quickConos(n.iterations = 10,
           names(panel),
           n.cores.p2=40,
           n.cores.con=40, 
           cms = unname(panel))

con <- buildConosGraph(con=cone,
                           k.conos=15, 
                           k.self=30, 
                           space="PCA", 
                           ncomps=40, 
                           n.odgenes=3e3, 
                           matching.method="mNN", 
                           metric="angular", 
                           score.component.variance=T,
                           alignment.strength=0,
                           min.dist=0.01, 
                           spread=14,
                           n.iterations=10)

con$plotGraph()
```


The cell clusters were annotated manually by plotting some marker genes.
We will not rerun the whole conos pipeline for this, we will load the objects now and show the marker genes used for annotation:

These notebooks are not run for this preview again because it takes a long time to load the data, filter it and process.


inhibitory neurons P60

```{r}
con60inh <- ci60$data.object
palinh60 <- ci60$cell.groups.palette
anno60inh <- ci60$cell.groups
```


```{r}
library(ggplot2)
con60inh$plotGraph(groups = anno60inh, font.size = 3) + scale_color_manual(values = palinh60) + theme_void() + theme(legend.position = "none")
```
```{r}
library(dplyr)
gs <- ci60$cell.groups %>% unique %>% as.character %>% strsplit(., split = "_") %>% unlist %>% unique

pls <- lapply(gs, function(x) con60inh$plotGraph(gene = x, plot.na = F, size = 0.5) + 
                theme_void() + ggtitle(x) + theme(legend.position = "none")
)
```

```{r, fig.height=10, fig.width=10}
cowplot::plot_grid(plotlist = pls)
```

```{r}

```

