---
title: "QC"
output: html_notebook
---

Author: Katarina Dragicevic (PhD student), Biotech Research and Innovation Centre, University of Copenhagen

This is an example flowork for cleaning both P15 and P60 data.
Both datasets were cleared for < 2000 UMI, < 0.05 mt genes and for doublets using scrublet. 
Both datasets had ambient RNA removed using Cellbender using the output plots for empty and total droplets from cell ranger. 
Some cells were additionally manually cleaned after running the pipeline multiple times because there were additional clusters which seemed to be doublets and low depth cells and which grouped together and messed up the plots.
The full vignette for CRMetrics can be found here: https://github.com/khodosevichlab/CRMetrics 


Loading packages:

```{r}
library(CRMetrics)
library(magrittr)
library(dplyr)
```


```{r}
crm <- CRMetrics$new(data.path = "/maps/projects/thor/data/allcounts15q/P15counts",
                     n.cores = 8, unique.names = F)
```

```{r}
metrics.to.plot <- crm$selectMetrics(ids = c(1:4,6,18,19))
crm$plotSummaryMetrics(metrics = metrics.to.plot,)
```

Some sample renaming

```{r}
currentnm <- c("P15_wt_AAM_k2"	,
"P15_wt_AAM_k1"	,
"P15_15q_AAM_k5",
"P15_15q_AAM_k4",
"P15_15q_AAM_k3",
"P15_15q_AAM_k2",
"P15_15q_AAM_k1",
"AAM3_P15wt",
"AAM12_15qb",
"AAM12_wtb"	,
"AAM12_15qa",
"AAM12_wta",
"AAM10_P15q15",
"AAM10_P15wt"	,
"AAM1_P15q15")

newnm <- c("P15_AAMK2_wt_b2",
"P15_AAMK1_wt_b2",
	"P15_AAMK5_15q_b2",
	"P15_AAMK4_15q_b2",
	"P15_AAMK3_15q_b2",
	"P15_AAMK2_15q_b2",
	"P15_AAMK1_15q_b2",
		"P15_AAM3_wt_b1",
	"P15_AAM12b_15q_b1",
"P15_AAM12b_wt_b1",
	"P15_AAM12a_15q_b1",
"P15_AAM12a_wt_b1",
	"P15_AAM10_15q_b1",
"P15_AAM10_wt_b1",
"P15_AAM1_15q_b1")


recodenm <- setNames(newnm, currentnm)
#crm$metadata$sample <- recode(crm$metadata$sample, !!!recodenm)
#crm$cms <- NULL
crm$addDetailedMetrics(unique.names = F)
cms <- crm$cms

cms2 <- lapply(cms, function(x) {
  cnm <- colnames(x) %>% strsplit(split = "!!", .) %>% sapply("[[", 2)
  colnames(x) <- cnm
  return(x)
  }
  )

names(cms2) <- recode(names(cms2), !!!recodenm)

cms3 <- lapply(names(cms2), function(x) {
  cms0 <- cms2[[x]]
  cnm <- paste0(x, "!!", colnames(cms0))
  colnames(cms0) <- cnm
  return(cms0)
  }
  ) %>% setNames(., names(cms2))



crm$cms <- cms3
crm$metadata$sample <- recode(crm$metadata$sample, !!!recodenm)
```


Metadata 

```{r}
cbind(Xist = crm$cms %>% lapply(function(x) x["Xist",] %>% sum),
      Eif2s3y = crm$cms %>% lapply(function(x) x["Eif2s3y",] %>% sum))

sex <- c("F", "F", "M", "M", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F")
batch <- c(rep("Batch1", 8), rep("Batch2", 7))
condition <- ifelse(grepl("15q", crm$cms %>% names), "15q", "wt")
metadata <- data.frame(sample = crm$cms %>% names, sex, batch, condition)

crm$metadata <- metadata
```

```{r}
crm$summary.metrics$sample <- recode(crm$summary.metrics$sample, !!!recodenm)
crm$detailed.metrics$sample <- recode(crm$detailed.metrics$sample, !!!recodenm)


crm$plotSummaryMetrics(comp.group = "condition",
                       metrics = metrics.to.plot, 
                       plot.geom = "point", 
                       stat.test = "non-parametric",
                       second.comp.group = "sex")

metrics.to.plot <- crm$detailed.metrics$metric %>%
  unique()
crm$plotDetailedMetrics(comp.group = "condition",
                        metrics = metrics.to.plot, 
                        plot.geom = "violin")


```

Run ambient RNA removal
```{r, fig.height=10, fig.width=10}
crm$prepareCellbender(shrinkage = 2, # Subsamples every 100th datapoint for faster plotting
                      show.expected.cells = TRUE, 
                      show.total.droplets = TRUE, n.cores = 10)

ec <- crm$getExpectedCells()
crm$expected.cells <- ec

td <- crm$getTotalDroplets()
crm$total.droplets <- td
```

```{r}
crm$expected.cells[[1]]
crm$total.droplets[[2]] <- 10000
crm$total.droplets[[3]] <- 25000
crm$total.droplets[[4]] <- 15000
crm$total.droplets[["AAM12_wta"]] <- 10000
crm$total.droplets[["AAM3_P15wt"]] <- 16000
crm$total.droplets[["P15_15q_AAM_k1"]] <- 30000
crm$total.droplets[["P15_15q_AAM_k2"]] <- 50000
crm$total.droplets[["P15_15q_AAM_k3"]] <- 40000
crm$total.droplets[["P15_15q_AAM_k4"]] <- 27000
crm$total.droplets[["P15_15q_AAM_k5"]] <- 27000
crm$total.droplets[["P15_wt_AAM_k1"]] <- 45000
crm$expected.cells[["P15_15q_AAM_k1"]] <- 6000
crm$expected.cells[["P15_15q_AAM_k2"]] <- 9000
crm$expected.cells[["P15_15q_AAM_k3"]] <- 3500
crm$expected.cells[["P15_15q_AAM_k4"]] <- 6000
crm$expected.cells[["P15_15q_AAM_k5"]] <- 6500
crm$expected.cells[["P15_wt_AAM_k1"]] <- 7000
crm$expected.cells[["P15_wt_AAM_k2"]] <- 7000

crm$prepareCellbender(shrinkage = 100, # Subsamples every 100th datapoint for faster plotting
                      show.expected.cells = TRUE, 
                      show.total.droplets = TRUE, n.cores = 10, expected.cells = crm$expected.cells, total.droplets = crm$total.droplets)
```


```{r}
crm$saveCellbenderScript(file = "cellbender_scriptP15.sh", 
                         fpr = 0.01, 
                         epochs = 120, 
                         use.gpu = TRUE,
                         expected.cells = crm$expected.cells, total.droplets = crm$total.droplets)
library(qs)
qsave(crm, "crm_p15.qs")
```


```{r}
crm <- qread("crm_p15.qs")
crm$addDetailedMetrics(cellbender = T,n.cores = 100)
crm$plotCbCellProbs()
crm$plotCbTraining()
crm$plotCbAmbExp(cutoff = 0.005)
crm$plotCbAmbGenes(cutoff = 0.005)
crm$plotCbCells()
```



Running pagoda2 + conos preview for depth cutoff

```{r}
crm$doPreprocessing(n.cores = 100)
crm$createEmbedding(n.cores = 100)
crm$plotEmbedding()
crm$plotEmbedding(depth = TRUE, 
             depth.cutoff = 1e3)
```


```{r}
crm$plotEmbedding(mito.frac = TRUE, 
             mito.cutoff = 0.05, 
             species = "mouse")
```

```{r}
crm$detectDoublets(env = "r-reticulate-dd",
                   conda.path = "/opt/software/miniconda/4.12.0/condabin/conda",
                   method = "scrublet")
crm$plotEmbedding(doublet.scores = T, doublet.method = "scrublet")


crm$plotFilteredCells(type = "embedding", 
                      depth = TRUE, 
                      depth.cutoff = 2000, 
                      doublet.method = "scrublet", 
                      mito.frac = TRUE, 
                      mito.cutoff = 0.05, 
                      species = "mouse")

crm$plotFilteredCells(type = "tile", 
                      doublet.method = "scrublet",
                      depth = TRUE, 
                      depth.cutoff = 2000,
                      mito.frac = TRUE, 
                      mito.cutoff = 0.05, 
                      species = "mouse")

crm$plotDepth(cutoff = 2000) 
```


```{r}
crm$filterCms(depth.cutoff = 2000, 
              mito.cutoff = 0.05, 
              doublets = "scrublet",
              samples.to.exclude = NULL,
              species = "mouse")
qsave(crm$cms.filtered, "cms_filtered_P15.qs", 
      nthreads = 15)
qsave(crm, "crm_P15.qs", 
      nthreads = 15)
```

